---
title: "CompMusic Porftolio"
author: "Natalia Zawada"
date: "21/02/2021"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
    orientation: columns
    storyboard: true
    theme: lumen
---

```{r}
library(tidyverse)
library(spotifyr)
library(flexdashboard)
library(ggthemes)
library(ggplot2)
library(plotly)
library(dplyr)


```

Introduction
=========================================

>Please do not show this portfolio in class!

I will be exploring the genre of techno and house. 
Techno emerged in the late 80's from the wide availability of synthesizers and the inspiration of early electronic music artist like Kraftwerk, Giorgio Mordor, Depech Mode, Front 242. 
The techno seed sprouted at about the same time in America and Europe. 
Chicago was the home of house and acid house, centering around Trax Records.
Detriot was the home of techno with many founding fathers including Juan Atkins, Derrick May, Jeff Mills. They owned MetroPlex, Transmat Records and UR Records respectively. 
Techno and house in Europe had its strong beginnings in Germany with hot pockets in Belguim and UK. Tresor Records in Berlin and R&S Records in Ghent egnited the techno flame. 

My corpus, for now, consists of three playlists - Detriot, Chicago, Europe. I looked at the record releases from the above mentioned labels, from first releases until about 1993, and searched for available tracks on Spotify. 

From its inception until early 1990's, there is clear geographical comparison to look at - Detriot Techno vs Berlin Techno. After 1993 or so, many Detriot founders move to Berlin... 


I would also like to look at ...  

```{r, echo=FALSE}

my_id <- 'natzaw'


detriot <- get_playlist_audio_features(my_id, "7zJxmBOEVValZYC0zmfLrf")
chicago <- get_playlist_audio_features(my_id, "5ZixL1Pb2mzyED9hSQ8y2G")
europe  <- get_playlist_audio_features(my_id, "1YnOhBchUwaRFOqzWZp4CV")

```
Visualizations {.storyboard}
=========================================

### Danceability vs Energy 

```{r, echo=FALSE}

ggplot(detriot, (aes(energy, danceability, color = key, size = tempo ))) +
  geom_point() +
  ggtitle("Detriot")

ggplot(chicago, (aes(energy, danceability, color = key, size = tempo))) +
  geom_point() + 
  ggtitle("Chicago")

ggplot(europe, (aes(energy, danceability, color = key, size = tempo))) +
  geom_point() + 
  ggtitle("Berlin")
```

***

Commentary Here

### Valence vs Loudness

```{r, echo=FALSE}
ggplot(detriot, (aes(valence, loudness, color = mode))) +
  geom_point() +
  ggtitle("Detriot")

ggplot(chicago, (aes(valence, loudness, color = mode))) +
  geom_point() + 
  ggtitle("Chicago")

ggplot(europe, (aes(valence, loudness, color = mode))) +
  geom_point() + 
  ggtitle("Berlin")
```

***

Commentary Here

### Chart 3

```{r, echo=FALSE}

```

***

Commentary Here


Discussion
=========================================

Discussion Here




